\begin{abstract}

  \par In this paper we outline an implementation of an octree complete with a raytracer that parses complex objects and renders them using octree traversal. We wrote a complete octree class from scratch with the help of some online resources and additionally wrote an object parser that incorporates the parser from our ray tracer project. Ultimately, this project stands as a large extension of the midterm raytracer in its goal to render more complex objects and speed up the rendering. 
  \par Unfortunately our octree implementation, though seemingly thorough and complete, is buggy. It works and gives significant speedups but is finicky and sometimes renders incomplete images. Thus only our object parser remains as a fully functioning element, it runs extremely slowly (highlighting the need for our octree implementation). We believe our octree class itself to be almost completely correct and useable, however our traversal algorithms/methods are likely incorrect somewhere as this seems to be where shapes are being dropped and not rendered. In the end it came down to time but we could not figure out how to fix the octree completely despite endless debugging. Despite this, we will go into extreme detail outlining the methods and ideas behind our traversal algorithms as we believe they only falter in our implementation.

  
%In this paper, we describe project Caribou, an augmented reality environment the models herd behavior. 

%The abstract should be in the neighborhood of 150-300 words total.  It should
%summarize the entire paper.  Write this after you have everything else, so that
%you know what to summarize.

\end{abstract}
