\section{Conclusion}

\par Overall the project was not a complete failure. We created what we believe to be an almost fully functional octree which in itself is an extraordinarily complicated structure to both create and to traverse. In addition we did end with one visible product– our object parser. 
\par We learned a lot from this trial and error, both about navigating 3D space and about designing a new data structure basically from scratch. A tremendous amount of thinking goes into simply thinking about organizing 3D space, let alone traversing it efficiently and this could easily be considered one of the biggest gains of the project, especially considering we were ultimately unable to get it to work.
\par This project should stand testament to the unfortunate reality that creating new things is super hard; not only is it hard starting a data structure from scratch, but even harder to do is start a relatively undocumented data structure from scratch is far worse. Finding documentation on octrees and how to traverse them effectively was pretty difficult, which was shocking as we thought that it would likely be a commonly used structure. As it turns out it appears that octrees are a somewhat outdated structure that has been replaced both by other similar structures like KD-Trees as well as parallel programming artifacts like CUDA.
\par Ultimately, if we were to continue working on our project in its current state we'd love to just continue debugging until we discovered what the issues are. Perhaps a misuse of a vital equation, or perhaps a simple missing line of code. We have no idea, but we wanted to see our baby work and that would easily be the goal for future work. As an even further extension we wished to test our octree on reflections and shadows as well, as both of those features required a search of the object list adding the octree would make it much faster (especially reflections as they take forever). Perhaps we could even try adding the Stanford bunny to a mirror room and see it in all its infinitely reflected glory.



%Briefly summarize what you did.

%What do you know now that you didn't at the time you started this project?
%What does your work tell us? What could we do next given the current status of the project?
